---
title: Stop detection in Vitaparcours movement data using segmentation algorithms
subtitle: A subtle subtitle
author: Judith Winet and Rodrigo Nogueira Silva
format:
  html:
    code-fold: true
execute:
  warning: false
  message: false
lang: en  # switch to "de" if you write your report in german
bibliography: bibliography.bib
---

```{r preprocessing}
#| code-summary: preprocessing

library("sf")
library("dplyr")
library("purrr")
library("slider")
library("ggplot2")
library("patchwork")
library("igraph")
library("kableExtra")

# loading the cleand data with static events with different time windows (from preprocessing file)

# 10 Seconds Window
Strava_1_processed_10    <- st_read("datasets/cleaned/static/Strava_1_processed_10.gpkg", quiet = TRUE)
Strava_2_processed_10    <- st_read("datasets/cleaned/static/Strava_2_processed_10.gpkg", quiet = TRUE)
Swisstopo_1_processed_10 <- st_read("datasets/cleaned/static/Swisstopo_1_processed_10.gpkg", quiet = TRUE)
Swisstopo_2_processed_10 <- st_read("datasets/cleaned/static/Swisstopo_2_processed_10.gpkg", quiet = TRUE)
Posmo_1_processed_10     <- st_read("datasets/cleaned/static/Posmo_1_processed_10.gpkg", quiet = TRUE)
Posmo_2_processed_10     <- st_read("datasets/cleaned/static/Posmo_2_processed_10.gpkg", quiet = TRUE)

# 20 Seconds Window
Strava_1_processed_20    <- st_read("datasets/cleaned/static/Strava_1_processed_20.gpkg", quiet = TRUE)
Strava_2_processed_20    <- st_read("datasets/cleaned/static/Strava_2_processed_20.gpkg", quiet = TRUE)
Swisstopo_1_processed_20 <- st_read("datasets/cleaned/static/Swisstopo_1_processed_20.gpkg", quiet = TRUE)
Swisstopo_2_processed_20 <- st_read("datasets/cleaned/static/Swisstopo_2_processed_20.gpkg", quiet = TRUE)
Posmo_1_processed_20     <- st_read("datasets/cleaned/static/Posmo_1_processed_20.gpkg", quiet = TRUE)
Posmo_2_processed_20     <- st_read("datasets/cleaned/static/Posmo_2_processed_20.gpkg", quiet = TRUE)

# 30 Seconds Window
Strava_1_processed_30    <- st_read("datasets/cleaned/static/Strava_1_processed_30.gpkg", quiet = TRUE)
Strava_2_processed_30    <- st_read("datasets/cleaned/static/Strava_2_processed_30.gpkg", quiet = TRUE)
Swisstopo_1_processed_30 <- st_read("datasets/cleaned/static/Swisstopo_1_processed_30.gpkg", quiet = TRUE)
Swisstopo_2_processed_30 <- st_read("datasets/cleaned/static/Swisstopo_2_processed_30.gpkg", quiet = TRUE)
Posmo_1_processed_30     <- st_read("datasets/cleaned/static/Posmo_1_processed_30.gpkg", quiet = TRUE)
Posmo_2_processed_30     <- st_read("datasets/cleaned/static/Posmo_2_processed_30.gpkg", quiet = TRUE)

# 40 Seconds Window
Strava_1_processed_40    <- st_read("datasets/cleaned/static/Strava_1_processed_40.gpkg", quiet = TRUE)
Strava_2_processed_40    <- st_read("datasets/cleaned/static/Strava_2_processed_40.gpkg", quiet = TRUE)
Swisstopo_1_processed_40 <- st_read("datasets/cleaned/static/Swisstopo_1_processed_40.gpkg", quiet = TRUE)
Swisstopo_2_processed_40 <- st_read("datasets/cleaned/static/Swisstopo_2_processed_40.gpkg", quiet = TRUE)
Posmo_1_processed_40     <- st_read("datasets/cleaned/static/Posmo_1_processed_40.gpkg", quiet = TRUE)
Posmo_2_processed_40     <- st_read("datasets/cleaned/static/Posmo_2_processed_40.gpkg", quiet = TRUE)

# way points cleaned

wp_olten_clean    <- st_read("datasets/cleaned/Waypoints_Olten.gpkg", quiet = TRUE)
wp_wollerau_clean <- st_read("datasets/cleaned/Waypoints_Wollerau.gpkg", quiet = TRUE)

forest_wollerau <- st_read("datasets/wald_selection_wollerau.gpkg", quiet = TRUE)

``` 

```{r Clustering and comparing to different buffers}
#| code-summary: Clustering and comparing to different buffers

# clusters together points which are close together and compare them to the way point buffer

stops <- function(full_df, waypoints, max_dist_list, wp_buffer_list) {
  
  # only take static points
  static_only <- full_df |> 
    filter(is_static == TRUE)
  
  # Checks if there are no static points
  if (nrow(static_only) == 0) {
    return(st_as_sf(data.frame(geometry = st_sfc(), crs = st_crs(waypoints))))
  }
  
  all_results_list <- list()
  
  # cluster for each maximale distanz 
  for (md in max_dist_list) {
    
    # calculates distance matrix for all points
    # distance and adjacency matrix for the current max_dist (md)
    dist_matrix <- st_distance(static_only)
    adj_matrix <- dist_matrix <= units::set_units(md, "m")
    
    # give Cluster-IDs 
    g <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "undirected")
    static_only$cluster_id <- igraph::components(g)$membership
    
    # calculate centroids for specific max_dist
    current_centroids <- static_only |>
      group_by(cluster_id) |>
      summarise(.groups = "drop") |>
      st_centroid() |>
      mutate(max_dist_setting = md) # saves maximum distance at which this centroid was calculated
    
    # inner loop: For each buffer size around the waypoints
    for (buf_size in wp_buffer_list) {
      
      # calculate temporary buffer for the waypoints
      temp_buffer_df <- waypoints |>
        mutate(gt_id = row_number()) |>
        st_buffer(buf_size)
      
      intersect_list <- st_intersects(current_centroids, temp_buffer_df)
      
      # generate dynamic column names (e.g. in_buf_md10_b15)
      col_name_bool <- paste0("in_buf_md", md, "_b", buf_size)
      col_name_id   <- paste0("gt_id_md", md, "_b", buf_size)
      
      # Write the results as new columns in the temporary Centroid table
      current_centroids <- current_centroids |>
        mutate(
          !!col_name_bool := lengths(intersect_list) > 0,
          !!col_name_id   := sapply(intersect_list, function(x) if(length(x) > 0) temp_buffer_df$gt_id[x[1]] else NA)
        )
    }
    
    anchor_id_col <- paste0("gt_id_md", md, "_b", max(wp_buffer_list))
    
    hits   <- current_centroids |> filter(!is.na(.data[[anchor_id_col]]))
    misses <- current_centroids |> filter(is.na(.data[[anchor_id_col]]))
    
    if (nrow(hits) > 0) {
      # Build precise column tracking strings matching the loop state
      all_dynamic_cols <- c(
        paste0("in_buf_md", md, "_b", wp_buffer_list),
        paste0("gt_id_md", md, "_b", wp_buffer_list)
      )
      
      # Squash multiple tracking dots in the same physical buffer down to 1 row
      hits_cleaned <- hits |>
        group_by(.data[[anchor_id_col]]) |>
        summarise(
          cluster_id       = min(cluster_id),
          max_dist_setting = min(max_dist_setting),
          # any_of safely skips any columns that were omitted
          across(any_of(all_dynamic_cols), ~ max(.x, na.rm = TRUE)),
          .groups          = "drop"
        ) |>
        st_centroid() # Re-compute single center geometry
      
      current_centroids <- bind_rows(hits_cleaned, misses)
    }
    
    # Save the interim result for this max_dist in our list.
    all_results_list[[as.character(md)]] <- current_centroids
  }
  
  # Combine all the partial results (for each max_dist) into one large table.
  final_df <- bind_rows(all_results_list)
  
  return(final_df)
}

# buffer size around way points
max_dist_liste <- c(5, 10, 15)
wp_buffer_liste <- c(5, 10, 15)

# add stops to 10 second windows

Strava_1_10_stops_list <- stops(Strava_1_processed_10, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Strava_2_10_stops_list <- stops(Strava_2_processed_10, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_1_10_stops_list <- stops(Swisstopo_1_processed_10, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_2_10_stops_list <- stops(Swisstopo_2_processed_10, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_1_10_stops_list <- stops(Posmo_1_processed_10, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_2_10_stops_list <- stops(Posmo_2_processed_10, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)

# add stops to 20 second windows

Strava_1_20_stops_list <- stops(Strava_1_processed_20, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Strava_2_20_stops_list <- stops(Strava_2_processed_20, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_1_20_stops_list <- stops(Swisstopo_1_processed_20, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_2_20_stops_list <- stops(Swisstopo_2_processed_20, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_1_20_stops_list <- stops(Posmo_1_processed_20, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_2_20_stops_list <- stops(Posmo_2_processed_20, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)

# add stops to 30 second windows

Strava_1_30_stops_list <- stops(Strava_1_processed_30, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Strava_2_30_stops_list <- stops(Strava_2_processed_30, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_1_30_stops_list <- stops(Swisstopo_1_processed_30, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_2_30_stops_list <- stops(Swisstopo_2_processed_30, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_1_30_stops_list <- stops(Posmo_1_processed_30, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_2_30_stops_list <- stops(Posmo_2_processed_30, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)

# add stops to 30 second windows

Strava_1_40_stops_list <- stops(Strava_1_processed_40, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Strava_2_40_stops_list <- stops(Strava_2_processed_40, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_1_40_stops_list <- stops(Swisstopo_1_processed_40, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_2_40_stops_list <- stops(Swisstopo_2_processed_40, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_1_40_stops_list <- stops(Posmo_1_processed_40, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_2_40_stops_list <- stops(Posmo_2_processed_40, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)

```


## Abstract

## Introduction

@laube2011 investigate how temporal scale affects the calculation of movement parameters (speed, sinuosity and turning angle) of animal trajectories. 

## Material and Methods

### Data

The Data was collected with the three different mobile apps Strava, Swisstopo and PosmoProject between the 20. Mai 2026 and the 24. Mai 2026. The Strava data was recorded for a Vitaparcours track in Olten on two different days. The data from Swisstopo and PosmoProject on the other hand were recorded on a track in Wollerau at the same time and also on two different days. The Strava and Swisstopo data recorded every second a point whereas the PosmoProject app did not recode in a continuous interval. 
The ground truth point were recorded with the Swisstopo app as well because we could not find and datasets for the Vitaparcours online. The track in Olten consists of 14 stops and the track in Wollerau of 15.

### Preprocessing

The data from Strava, Swisstopo and PosmoProject was imported, standardised and converter to the Europ/Zurich time zone. Furthermore, all coordinate geometries were transformed into the Swiss coordinate systems CH1903+ / LV95, to ensure accurate calculations. To isolate static events from the different trajectories, geometries were classified using an adaptive mean temporal window distance for four different windows (10, 20, 30 and 40 seconds) (@laube2011). For each point i, the algorithem calcualted the neighbouring coordinates within the window ($[t_{i} - \frac{w}{2}, t_{i} + \frac{w}{2}]$) and computed their mean spatial distance (stepMean_i). The average of all these spatial means was then used as a threshold to calculate points as either static (is_static = TRUE) if the stepMean_i was smaller than the threshold. The resulting 24 datasets were then finally exported as GeoPackage files and used for further analysis. These processes were done in the preprocessing file as they were computationally very heavy (***need to add Link***).


## Results

```{r 10s Window}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Visualisation of the static events of all tracks with a time window of 10 seconds"
#| label: fig-10-window

# Plot the static events

olten_y <- seq(1244200, 1245100, by = 200)
olten_x <- seq(2633200, 2634000, by = 200)

wollerau_y <- seq(1226700, 1227200, by = 200)
wollerau_x <- seq(2694900, 2695700, by = 200)

base_theme <- list(
  theme_minimal(),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5, size = 8),
    axis.text.y = element_text(size = 8),
    plot.title = element_text(size = 11, face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank() # Removes messy minor gridlines
  )
)

# hide legend for the first 5 plots to prevent duplicates
hide_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  guide = "none"
)

# show full legend on the very last plot, forcing all 3 colors to display
show_full_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  limits = c(TRUE, FALSE, NA), # <--- Keep this order
  labels = c("Static", "Moving", "Unknown/Gap"), # <--- Match the order of limits exactly without names
  name = "Status"
)

p1 <- ggplot(Strava_1_processed_10 ) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 1") +
  scale_x_continuous(breaks = olten_x) + 
  scale_y_continuous(breaks = olten_y) +
  coord_sf(datum = 2056) + 
  base_theme

p2 <- ggplot(Strava_2_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 2") +
  scale_x_continuous(breaks = olten_x) + 
  scale_y_continuous(breaks = olten_y) +
  coord_sf(datum = 2056) + 
  base_theme

p3 <- ggplot(Swisstopo_1_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 1") +
  scale_x_continuous(breaks = wollerau_x) + 
  scale_y_continuous(breaks = wollerau_y) +
  coord_sf(datum = 2056) + 
  base_theme

p4 <- ggplot(Swisstopo_2_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 2") +
  scale_x_continuous(breaks = wollerau_x) + 
  scale_y_continuous(breaks = wollerau_y) +
  coord_sf(datum = 2056) + 
  base_theme

p5 <- ggplot(Posmo_1_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Posmo 1") +
  scale_x_continuous(breaks = wollerau_x) + 
  scale_y_continuous(breaks = wollerau_y) +
  coord_sf(datum = 2056) + 
  base_theme

p6 <- ggplot(Posmo_2_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  show_full_legend +
  labs(title = "Posmo 2") +
  scale_x_continuous(breaks = wollerau_x) + 
  scale_y_continuous(breaks = wollerau_y) +
  coord_sf(datum = 2056) + 
  base_theme

# creates a grid layout with 3 columns, and collects the legends into one clean spot
p1 + p2 + p3 + p4 + p5 + p6 + 
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Static vs. Moving",
    subtitle = "Time window of 10 seconds",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  ) 

```

```{r 20s Window}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Visualisation of the static events of all tracks with a time window of 20 seconds"
#| label: fig-20-window

# Plot the static events

theme_x_cord <- list(
  theme_minimal(),
  coord_sf(datum = 2056),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
    plot.title = element_text(size = 10, face = "bold")
  )
)

# hide legend for the first 5 plots to prevent duplicates
hide_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  guide = "none"
)

# show full legend on the very last plot, forcing all 3 colors to display
show_full_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  limits = c(TRUE, FALSE, NA), # <--- Keep this order
  labels = c("Static", "Moving", "Unknown/Gap"), # <--- Match the order of limits exactly without names
  name = "Status"
)

p1 <- ggplot(Strava_1_processed_20 ) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 1") +
  theme_x_cord

p2 <- ggplot(Strava_2_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 2") +
  theme_x_cord

p3 <- ggplot(Swisstopo_1_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 1") +
  theme_x_cord

p4 <- ggplot(Swisstopo_2_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 2") +
  theme_x_cord

p5 <- ggplot(Posmo_1_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Posmo 1") +
  theme_x_cord

p6 <- ggplot(Posmo_2_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  show_full_legend +
  labs(title = "Posmo 2") +
  theme_x_cord

# creates a grid layout with 3 columns, and collects the legends into one clean spot
p1 + p2 + p3 + p4 + p5 + p6 + 
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Static vs. Moving",
    subtitle = "Time window of 20 seconds",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  ) 

```

```{r 30s Window}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Visualisation of the static events of all tracks with a time window of 30 seconds"
#| label: fig-30-window

# Plot the static events

theme_x_cord <- list(
  theme_minimal(),
  coord_sf(datum = 2056),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
    plot.title = element_text(size = 10, face = "bold")
  )
)

# hide legend for the first 5 plots to prevent duplicates
hide_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  guide = "none"
)

# show full legend on the very last plot, forcing all 3 colors to display
show_full_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  limits = c(TRUE, FALSE, NA), # <--- Keep this order
  labels = c("Static", "Moving", "Unknown/Gap"), # <--- Match the order of limits exactly without names
  name = "Status"
)

p1 <- ggplot(Strava_1_processed_30 ) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 1") +
  theme_x_cord

p2 <- ggplot(Strava_2_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 2") +
  theme_x_cord

p3 <- ggplot(Swisstopo_1_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 1") +
  theme_x_cord

p4 <- ggplot(Swisstopo_2_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 2") +
  theme_x_cord

p5 <- ggplot(Posmo_1_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Posmo 1") +
  theme_x_cord

p6 <- ggplot(Posmo_2_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  show_full_legend +
  labs(title = "Posmo 2") +
  theme_x_cord

# creates a grid layout with 3 columns, and collects the legends into one clean spot
p1 + p2 + p3 + p4 + p5 + p6 + 
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Static vs. Moving",
    subtitle = "Time window of 20 seconds",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  ) 

```

```{r 40s Window}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Visualisation of the static events of all tracks with a time window of 40 seconds"
#| label: fig-40-window

# Plot the static events

theme_x_cord <- list(
  theme_minimal(),
  coord_sf(datum = 2056),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
    plot.title = element_text(size = 10, face = "bold")
  )
)

# hide legend for the first 5 plots to prevent duplicates
hide_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  guide = "none"
)

# show full legend on the very last plot, forcing all 3 colors to display
show_full_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  limits = c(TRUE, FALSE, NA), # <--- Keep this order
  labels = c("Static", "Moving", "Unknown/Gap"), # <--- Match the order of limits exactly without names
  name = "Status"
)

p1 <- ggplot(Strava_1_processed_40 ) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 1") +
  theme_x_cord

p2 <- ggplot(Strava_2_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 2") +
  theme_x_cord

p3 <- ggplot(Swisstopo_1_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 1") +
  theme_x_cord

p4 <- ggplot(Swisstopo_2_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 2") +
  theme_x_cord

p5 <- ggplot(Posmo_1_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Posmo 1") +
  theme_x_cord

p6 <- ggplot(Posmo_2_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  show_full_legend +
  labs(title = "Posmo 2") +
  theme_x_cord

# creates a grid layout with 3 columns, and collects the legends into one clean spot
p1 + p2 + p3 + p4 + p5 + p6 + 
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Static vs. Moving",
    subtitle = "Time window of 40 seconds",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  ) 

```

```{r Example: Posmo 1 30s Window}
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Visualisation of the static events (with a time window of 30 seconds) inside the buffer for different cluster distances and buffer sizes for Posmo 1"
#| label: fig-30s-posmo1


# plot the combinations of clustering and buffering around waypoint

theme_x_cord <- list(
  coord_sf(datum = 2056),
  theme_minimal(),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
    plot.title = element_text(size = 9, face = "bold"),
    axis.title = element_blank()
  )
)

plot_liste <- list()

for (md in max_dist_liste) {
  
  filtered <- Posmo_1_30_stops_list |> 
    filter(max_dist_setting == md)
  
  for (buf in wp_buffer_liste) {
    
    current_col <- paste0("in_buf_md", md, "_b", buf)
    
    p <- ggplot() +
      geom_sf(data = st_buffer(wp_wollerau_clean, buf), 
              fill = "gray70", 
              alpha = 0.50, 
              color = NA) +
      geom_sf(data = filtered, 
              aes(color = as.character(.data[[current_col]])), 
              size = 1) +
      scale_color_manual(
        values = c("1" = "darkgreen", "0" = "blue"),
        labels = c("1" = "Hit (TP)", "0" = "Outside of Buffer (FP)"),
        drop = FALSE
      ) +
      labs(title = paste0("MD: ", md, "m | Buffer: ", buf, "m"), 
           color = "Inside of the buffer") +
      theme_x_cord
    
    plot_liste[[paste0("md", md, "_b", buf)]] <- p
  }
}

wrap_plots(plot_liste, ncol = 3) + 
  plot_layout(guides = "collect") + 
  plot_annotation(
    title = "Posmo 1 - 9 combinations of clustering and buffering around waypoints",
    subtitle = "Temporal window for static events = 30 seconds | Cluster-Distanz (MD) | Waypoint-Buffer (Buffer)"
  )
```

```{r Parameter Table with Highlighting}
#| tbl-cap: "Parameter Table evaluating out of maximum 88 possible ground truth waypoint targets across all combined tracking assets. The top 5 performing rows are highlighted in Green and the bottom 5 in Red."
#| label: tbl-parameters-all-prov



# 1. Safe helper functions for dynamic column evaluation
extract_sum <- function(df, md, buf, val) {
  col_name <- paste0("in_buf_md", md, "_b", buf)
  if (!col_name %in% colnames(df)) return(0)
  sum(df[[col_name]] == val, na.rm = TRUE)
}

extract_distinct <- function(df, md, buf) {
  col_name <- paste0("gt_id_md", md, "_b", buf)
  if (!col_name %in% colnames(df)) return(0)
  dplyr::n_distinct(df[[col_name]], na.rm = TRUE)
}

# 2. Master function to process a dataset
get_all_permutations <- function(df, app_name, track_id, window_sec, total_gt) {
  if (nrow(df) == 0) return(data.frame())
  distances <- c(5, 10, 15)
  buffers   <- c(5, 10, 15)
  param_grid <- expand.grid(Max_Dist = distances, WP_Buffer = buffers)
  
  map_dfr(1:nrow(param_grid), function(i) {
    md  <- param_grid$Max_Dist[i]
    buf <- param_grid$WP_Buffer[i]
    df_sub <- df |> filter(max_dist_setting == md)
    if (nrow(df_sub) == 0) return(data.frame())
    
    data.frame(
      App         = app_name,
      Track       = track_id,
      Window      = paste0(window_sec, "s"),
      Max_Dist    = md,
      WP_Buffer   = buf,
      Total_Stops = nrow(df_sub),
      True_Pos    = extract_sum(df_sub, md, buf, TRUE),
      False_Pos   = extract_sum(df_sub, md, buf, FALSE),
      Captured    = extract_distinct(df_sub, md, buf),
      Total_GT    = total_gt
    )
  })
}

# 3. Explicit ground truth variables
gt_strava_t1 <- nrow(wp_olten_clean)    # 14
gt_strava_t2 <- nrow(wp_olten_clean)    # 14
gt_swiss_t1  <- nrow(wp_wollerau_clean) # 15
gt_swiss_t2  <- nrow(wp_wollerau_clean) # 15
gt_posmo_t1  <- nrow(wp_wollerau_clean) # 15
gt_posmo_t2  <- nrow(wp_wollerau_clean) # 15
# Maximum possible waypoints to detect across all 6 tracks combined = 88
MAX_GLOBAL_GT <- (14 * 2) + (15 * 4) 

# 4. Run calculations over the 24 clean datasets
master_raw_metrics <- bind_rows(
  get_all_permutations(Strava_1_10_stops_list, "Strava", "T1", 10, gt_strava_t1),
  get_all_permutations(Strava_2_10_stops_list, "Strava", "T2", 10, gt_strava_t2),
  get_all_permutations(Swisstopo_1_10_stops_list, "Swisstopo", "T1", 10, gt_swiss_t1),
  get_all_permutations(Swisstopo_2_10_stops_list, "Swisstopo", "T2", 10, gt_swiss_t2),
  get_all_permutations(Posmo_1_10_stops_list, "Posmo", "T1", 10, gt_posmo_t1),
  get_all_permutations(Posmo_2_10_stops_list, "Posmo", "T2", 10, gt_posmo_t2),
  
  get_all_permutations(Strava_1_20_stops_list, "Strava", "T1", 20, gt_strava_t1),
  get_all_permutations(Strava_2_20_stops_list, "Strava", "T2", 20, gt_strava_t2),
  get_all_permutations(Swisstopo_1_20_stops_list, "Swisstopo", "T1", 20, gt_swiss_t1),
  get_all_permutations(Swisstopo_2_20_stops_list, "Swisstopo", "T2", 20, gt_swiss_t2),
  get_all_permutations(Posmo_1_20_stops_list, "Posmo", "T1", 20, gt_posmo_t1),
  get_all_permutations(Posmo_2_20_stops_list, "Posmo", "T2", 20, gt_posmo_t2),
  
  get_all_permutations(Strava_1_30_stops_list, "Strava", "T1", 30, gt_strava_t1),
  get_all_permutations(Strava_2_30_stops_list, "Strava", "T2", 30, gt_strava_t2),
  get_all_permutations(Swisstopo_1_30_stops_list, "Swisstopo", "T1", 30, gt_swiss_t1),
  get_all_permutations(Swisstopo_2_30_stops_list, "Swisstopo", "T2", 30, gt_swiss_t2),
  get_all_permutations(Posmo_1_30_stops_list, "Posmo", "T1", 30, gt_posmo_t1),
  get_all_permutations(Posmo_2_30_stops_list, "Posmo", "T2", 30, gt_posmo_t2),
  
  get_all_permutations(Strava_1_40_stops_list, "Strava", "T1", 40, gt_strava_t1),
  get_all_permutations(Strava_2_40_stops_list, "Strava", "T2", 40, gt_strava_t2),
  get_all_permutations(Swisstopo_1_40_stops_list, "Swisstopo", "T1", 40, gt_swiss_t1),
  get_all_permutations(Swisstopo_2_40_stops_list, "Swisstopo", "T2", 40, gt_swiss_t2),
  get_all_permutations(Posmo_1_40_stops_list, "Posmo", "T1", 40, gt_posmo_t1),
  get_all_permutations(Posmo_2_40_stops_list, "Posmo", "T2", 40, gt_posmo_t2)
)

# 5. Corrected Mathematical Aggregation
global_parameter_table <- master_raw_metrics |>
  group_by(Window, Max_Dist, WP_Buffer) |>
  summarise(
    Total_Stops_Found  = sum(Total_Stops),
    True_Positives     = sum(True_Pos),
    False_Positives    = sum(False_Pos),
    
    # Core Fixes:
    # Waypoints Successfully Detected (capped strictly at max 88)
    Waypoints_Hit      = sum(Captured),
    # False Negatives = Total available ground truth points (88) minus the ones we successfully hit
    False_Negatives    = MAX_GLOBAL_GT - Waypoints_Hit,
    # Detection Rate = What % of the 88 locations did the algorithm find
    Avg_Detection_Rate = round((Waypoints_Hit / MAX_GLOBAL_GT) * 100, 1),
    .groups = "drop"
  ) |>
  arrange(Window, Max_Dist, WP_Buffer)

# 6. Rank performances contextually behind the scenes
ranked_indices <- global_parameter_table |>
  mutate(row_id = row_number()) |>
  arrange(desc(Avg_Detection_Rate), False_Positives)

top_5_rows    <- head(ranked_indices$row_id, 5)
bottom_5_rows <- tail(ranked_indices$row_id, 5)

# 7. Render Styled Chronological Table
global_parameter_table |>
  select(Window, Max_Dist, WP_Buffer, Total_Stops_Found, True_Positives, False_Positives, False_Negatives, Avg_Detection_Rate) |>
  knitr::kable(
    format = "html",
    col.names = c("Time Window", "Clustering Max Dist (m)", "Waypoint Buffer (m)", "Total Stops Found", "True Positives (TP)", "False Positives (FP)", "False Negatives (FN)", "Stops detected inside buffer (%)")
  ) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  row_spec(top_5_rows, background = "#d4edda", extra_css = "font-weight: bold; color: #155724;") |>
  row_spec(bottom_5_rows, background = "#f8d7da", extra_css = "color: #721c24;")
```


## Discussion

## Appendix

### AI use

### Wordcount

<!-- after installing the wordcountadding, remove the line "#| eval: false" -->

```{r}
# #| eval: false
# library("pacman")
# p_install_gh("benmarwick/wordcountaddin")

wordcountaddin::word_count("index.qmd")
```


