---
title: Stop detection in Vitaparcours movement data using segmentation algorithms
subtitle: A subtle subtitle
author: Judith Winet and Rodrigo Nogueira Silva
format:
  html:
    code-fold: true
execute:
  warning: false
  message: false
lang: en  # switch to "de" if you write your report in german
bibliography: bibliography.bib
---

```{r preprocessing}
#| code-summary: preprocessing

library("sf")
library("dplyr")
library("units")
library("purrr")
library("slider")
library("ggplot2")
library("patchwork")
library("igraph")
library("kableExtra")
library("ggplot2")

# loading the cleand data with static events with different time windows (from preprocessing file)

# 10 Seconds Window
Strava_1_processed_10    <- st_read("datasets/cleaned/static/Strava_1_processed_10.gpkg", quiet = TRUE)
Strava_2_processed_10    <- st_read("datasets/cleaned/static/Strava_2_processed_10.gpkg", quiet = TRUE)
Swisstopo_1_processed_10 <- st_read("datasets/cleaned/static/Swisstopo_1_processed_10.gpkg", quiet = TRUE)
Swisstopo_2_processed_10 <- st_read("datasets/cleaned/static/Swisstopo_2_processed_10.gpkg", quiet = TRUE)
Posmo_1_processed_10     <- st_read("datasets/cleaned/static/Posmo_1_processed_10.gpkg", quiet = TRUE)
Posmo_2_processed_10     <- st_read("datasets/cleaned/static/Posmo_2_processed_10.gpkg", quiet = TRUE)

# 20 Seconds Window
Strava_1_processed_20    <- st_read("datasets/cleaned/static/Strava_1_processed_20.gpkg", quiet = TRUE)
Strava_2_processed_20    <- st_read("datasets/cleaned/static/Strava_2_processed_20.gpkg", quiet = TRUE)
Swisstopo_1_processed_20 <- st_read("datasets/cleaned/static/Swisstopo_1_processed_20.gpkg", quiet = TRUE)
Swisstopo_2_processed_20 <- st_read("datasets/cleaned/static/Swisstopo_2_processed_20.gpkg", quiet = TRUE)
Posmo_1_processed_20     <- st_read("datasets/cleaned/static/Posmo_1_processed_20.gpkg", quiet = TRUE)
Posmo_2_processed_20     <- st_read("datasets/cleaned/static/Posmo_2_processed_20.gpkg", quiet = TRUE)

# 30 Seconds Window
Strava_1_processed_30    <- st_read("datasets/cleaned/static/Strava_1_processed_30.gpkg", quiet = TRUE)
Strava_2_processed_30    <- st_read("datasets/cleaned/static/Strava_2_processed_30.gpkg", quiet = TRUE)
Swisstopo_1_processed_30 <- st_read("datasets/cleaned/static/Swisstopo_1_processed_30.gpkg", quiet = TRUE)
Swisstopo_2_processed_30 <- st_read("datasets/cleaned/static/Swisstopo_2_processed_30.gpkg", quiet = TRUE)
Posmo_1_processed_30     <- st_read("datasets/cleaned/static/Posmo_1_processed_30.gpkg", quiet = TRUE)
Posmo_2_processed_30     <- st_read("datasets/cleaned/static/Posmo_2_processed_30.gpkg", quiet = TRUE)

# 40 Seconds Window
Strava_1_processed_40    <- st_read("datasets/cleaned/static/Strava_1_processed_40.gpkg", quiet = TRUE)
Strava_2_processed_40    <- st_read("datasets/cleaned/static/Strava_2_processed_40.gpkg", quiet = TRUE)
Swisstopo_1_processed_40 <- st_read("datasets/cleaned/static/Swisstopo_1_processed_40.gpkg", quiet = TRUE)
Swisstopo_2_processed_40 <- st_read("datasets/cleaned/static/Swisstopo_2_processed_40.gpkg", quiet = TRUE)
Posmo_1_processed_40     <- st_read("datasets/cleaned/static/Posmo_1_processed_40.gpkg", quiet = TRUE)
Posmo_2_processed_40     <- st_read("datasets/cleaned/static/Posmo_2_processed_40.gpkg", quiet = TRUE)

# way points cleaned

wp_olten_clean    <- st_read("datasets/cleaned/Waypoints_Olten.gpkg", quiet = TRUE)
wp_wollerau_clean <- st_read("datasets/cleaned/Waypoints_Wollerau.gpkg", quiet = TRUE)

forest_wollerau <- st_read("datasets/wald_selection_wollerau.gpkg", quiet = TRUE)

``` 

```{r Clustering and comparing to different buffers}
#| code-summary: Clustering and comparing to different buffers
#| lst-label: lst-cluster-buffer
#| lst-cap: Methods

# clusters together points which are close together and compare them to the way point buffer

stops <- function(full_df, waypoints, max_dist_list, wp_buffer_list) {
  
  # only take static points
  static_only <- full_df |> 
    filter(is_static == TRUE)
  
  # Checks if there are no static points
  if (nrow(static_only) == 0) {
    return(st_as_sf(data.frame(geometry = st_sfc(), crs = st_crs(waypoints))))
  }
  
  all_results_list <- list()
  
  # cluster for each maximale distanz 
  for (md in max_dist_list) {
    
    # calculates distance matrix for all points
    # distance and adjacency matrix for the current max_dist (md)
    dist_matrix <- st_distance(static_only)
    adj_matrix <- dist_matrix <= set_units(md, "m")
    
    # give Cluster-IDs 
    g <- graph_from_adjacency_matrix(adj_matrix, mode = "undirected")
    static_only$cluster_id <- components(g)$membership
    
    # calculate centroids for specific max_dist
    current_centroids <- static_only |>
      group_by(cluster_id) |>
      summarise(.groups = "drop") |>
      st_centroid() |>
      mutate(max_dist_setting = md) # saves maximum distance at which this centroid was calculated
    
    # inner loop: For each buffer size around the waypoints
    for (buf_size in wp_buffer_list) {
      
      # calculate temporary buffer for the waypoints
      temp_buffer_df <- waypoints |>
        mutate(gt_id = row_number()) |>
        st_buffer(buf_size)
      
      intersect_list <- st_intersects(current_centroids, temp_buffer_df)
      
      # generate dynamic column names (e.g. in_buf_md10_b15)
      col_name_bool <- paste0("in_buf_md", md, "_b", buf_size)
      col_name_id   <- paste0("gt_id_md", md, "_b", buf_size)
      
      # Write the results as new columns in the temporary Centroid table
      current_centroids <- current_centroids |>
        mutate(
          !!col_name_bool := lengths(intersect_list) > 0,
          !!col_name_id   := sapply(intersect_list, function(x) if(length(x) > 0) temp_buffer_df$gt_id[x[1]] else NA)
        )
    }
    
    anchor_id_col <- paste0("gt_id_md", md, "_b", max(wp_buffer_list))
    
    hits   <- current_centroids |> filter(!is.na(.data[[anchor_id_col]]))
    misses <- current_centroids |> filter(is.na(.data[[anchor_id_col]]))
    
    if (nrow(hits) > 0) {
      # Build precise column tracking strings matching the loop state
      all_dynamic_cols <- c(
        paste0("in_buf_md", md, "_b", wp_buffer_list),
        paste0("gt_id_md", md, "_b", wp_buffer_list)
      )
      
      # combine multiple tracking dots in the same physical buffer down to 1 row
      hits_cleaned <- hits |>
        group_by(.data[[anchor_id_col]]) |>
        summarise(
          cluster_id       = min(cluster_id),
          max_dist_setting = min(max_dist_setting),
          # any_of safely skips any columns that were omitted
          across(any_of(all_dynamic_cols), ~ max(.x, na.rm = TRUE)),
          .groups          = "drop"
        ) |>
        st_centroid() # Re-compute single center geometry
      
      current_centroids <- bind_rows(hits_cleaned, misses)
    }
    
    # Save the interim result for this max_dist in our list.
    all_results_list[[as.character(md)]] <- current_centroids
  }
  
  # Combine all the partial results (for each max_dist) into one large table.
  final_df <- bind_rows(all_results_list)
  
  return(final_df)
}

# buffer size around way points
max_dist_liste <- c(5, 10, 15)
wp_buffer_liste <- c(5, 10, 15)

# add stops to 10 second windows

Strava_1_10_stops_list <- stops(Strava_1_processed_10, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Strava_2_10_stops_list <- stops(Strava_2_processed_10, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_1_10_stops_list <- stops(Swisstopo_1_processed_10, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_2_10_stops_list <- stops(Swisstopo_2_processed_10, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_1_10_stops_list <- stops(Posmo_1_processed_10, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_2_10_stops_list <- stops(Posmo_2_processed_10, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)

# add stops to 20 second windows

Strava_1_20_stops_list <- stops(Strava_1_processed_20, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Strava_2_20_stops_list <- stops(Strava_2_processed_20, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_1_20_stops_list <- stops(Swisstopo_1_processed_20, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_2_20_stops_list <- stops(Swisstopo_2_processed_20, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_1_20_stops_list <- stops(Posmo_1_processed_20, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_2_20_stops_list <- stops(Posmo_2_processed_20, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)

# add stops to 30 second windows

Strava_1_30_stops_list <- stops(Strava_1_processed_30, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Strava_2_30_stops_list <- stops(Strava_2_processed_30, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_1_30_stops_list <- stops(Swisstopo_1_processed_30, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_2_30_stops_list <- stops(Swisstopo_2_processed_30, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_1_30_stops_list <- stops(Posmo_1_processed_30, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_2_30_stops_list <- stops(Posmo_2_processed_30, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)

# add stops to 40 second windows

Strava_1_40_stops_list <- stops(Strava_1_processed_40, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Strava_2_40_stops_list <- stops(Strava_2_processed_40, wp_olten_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_1_40_stops_list <- stops(Swisstopo_1_processed_40, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Swisstopo_2_40_stops_list <- stops(Swisstopo_2_processed_40, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_1_40_stops_list <- stops(Posmo_1_processed_40, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)
Posmo_2_40_stops_list <- stops(Posmo_2_processed_40, wp_wollerau_clean, max_dist = max_dist_liste, wp_buffer = wp_buffer_liste)

```


## Abstract

## Introduction

@laube2011 investigate how temporal scale affects the calculation of movement parameters (speed, sinuosity and turning angle) of animal trajectories. 

## Material and Methods

### Data

The data was collected with the three different mobile apps, Strava, Swisstopo and Posmo, between the 20th of May 2026 and the 24th of May 2026. The Strava data was recorded for a Vitaparcours track in Olten on two different days. The data from Swisstopo and Posmo, on the other hand, was recorded on a track in Wollerau at the same time and also on two different days. The Strava and Swisstopo data recorded every second a fix, whereas the Posmo app did not record at a continuous interval. 
The ground truth points were recorded with the Swisstopo app as well because we could not find any datasets for the Vitaparcours online. The track in Olten consists of 14 stops and the track in Wollerau of 15.

### Preprocessing

The data from Strava, Swisstopo and Posmo were imported, standardised and converted to the Europe/Zurich time zone. Furthermore, all coordinate geometries were transformed into the Swiss coordinate systems CH1903+ / LV95 to ensure accurate calculations (***need to add Link to Preprocessing file***). 

### Methods

#### Isolating static events

To isolate static events from the different trajectories, geometries were classified using an adaptive mean temporal window distance for four different windows (10, 20, 30 and 40 seconds) according to @laube2011. For each point i, the algorithm calculated the neighbouring coordinates within the window ($[t_{i} - \frac{w}{2}, t_{i} + \frac{w}{2}]$) and computed their mean spatial distance (stepMean_i). The average of all these spatial means was then used as a threshold to calculate points as static (is_static = TRUE) if the stepMean_i was smaller than the threshold. The resulting 24 datasets were then finally exported as GeoPackage files and used for further analysis (@fig-10-window, @fig-20-window, @fig-30-window and @fig-40-window). These processes were done in the preprocessing file, as they were computationally very heavy (***need to add Link to Preprocessing file***).

#### Clustering static events and comparing them to different buffers 

To identify and analyse single stationary events (stops) across the different tracking datasets, a spatial analysis workflow across multiple scales was implemented (@lst-cluster-buffer). This customised r function first filters static data points and applies a clustering mechanism based on distance to create one stop from multiple static fixes. To achieve this, points within a specific maximum distance threshold (5, 10 and 15 metres) are grouped together and their respective geometric centroids are calculated, resulting in one single stop per cluster.

Furthermore, a distance analysis evaluates these centroids against the ground truth waypoint locations. Multiple distance buffers (5, 10 and 15 metres) are generated around the waypoints to identify spatial intersections between the ground truth stop buffers and the generated stops. To avoid having multiple stops and thus multiple true positives per buffer, the stops are further filtered. Multiple classified stops inside the different buffers are combined into a single stop inside the buffer using the centroid. This ensures data integrity and makes data analysis and interpretation easier to understand.

This workflow dynamically tracks spatial attributes across different parameter combinations and generates structured spatial data frames. To further assess the temporal sensitivity, the entire analysis was repeated across the preprocessed datasets, which were segmented into four distinct temporal windows for capturing static events. This ensures a robust evaluation of spatial clustering behaviour across varying spatial and temporal scales.

#### Performance metrics Precision, Recall and F1-score 



## Results

@fig-10-window, @fig-20-window, @fig-30-window and @fig-40-window show the results of the static event identification for the four different temporal windows 10, 20, 30 and 40 seconds. Because the Posmo data was not captured in a continuous interval for some fixes in the temporal windows 10 and 20 seconds no neighbouring fixes were detected. These points are plotted as "Unknown" and could not be used for further analysis. 

```{r 10s Window}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Visualisation of the static events of all tracks with a time window of 10 seconds."
#| label: fig-10-window

# Plot the static events

olten_y <- seq(1244200, 1245100, by = 200)
olten_x <- seq(2633200, 2634000, by = 200)

wollerau_y <- seq(1226700, 1227200, by = 200)
wollerau_x <- seq(2694900, 2695700, by = 200)

base_theme <- list(
  theme_minimal(),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5, size = 8),
    axis.text.y = element_text(size = 8),
    plot.title = element_text(size = 11, face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank() # Removes messy minor gridlines
  )
)

# hide legend for the first 5 plots to prevent duplicates
hide_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  guide = "none"
)

# show full legend on the very last plot, forcing all 3 colors to display
show_full_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  limits = c(TRUE, FALSE, NA), # <--- Keep this order
  labels = c("Static", "Moving", "Unknown"), # <--- Match the order of limits exactly without names
  name = "Status"
)

p1 <- ggplot(Strava_1_processed_10 ) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 1") +
  scale_x_continuous(breaks = olten_x) + 
  scale_y_continuous(breaks = olten_y) +
  coord_sf(datum = 2056) + 
  base_theme

p2 <- ggplot(Strava_2_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 2") +
  scale_x_continuous(breaks = olten_x) + 
  scale_y_continuous(breaks = olten_y) +
  coord_sf(datum = 2056) + 
  base_theme

p3 <- ggplot(Swisstopo_1_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 1") +
  scale_x_continuous(breaks = wollerau_x) + 
  scale_y_continuous(breaks = wollerau_y) +
  coord_sf(datum = 2056) + 
  base_theme

p4 <- ggplot(Swisstopo_2_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 2") +
  scale_x_continuous(breaks = wollerau_x) + 
  scale_y_continuous(breaks = wollerau_y) +
  coord_sf(datum = 2056) + 
  base_theme

p5 <- ggplot(Posmo_1_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Posmo 1") +
  scale_x_continuous(breaks = wollerau_x) + 
  scale_y_continuous(breaks = wollerau_y) +
  coord_sf(datum = 2056) + 
  base_theme

p6 <- ggplot(Posmo_2_processed_10) +
  geom_sf(aes(color = is_static), size = 1) +
  show_full_legend +
  labs(title = "Posmo 2") +
  scale_x_continuous(breaks = wollerau_x) + 
  scale_y_continuous(breaks = wollerau_y) +
  coord_sf(datum = 2056) + 
  base_theme

# creates a grid layout with 3 columns, and collects the legends into one clean spot
p1 + p2 + p3 + p4 + p5 + p6 + 
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Static vs. Moving",
    subtitle = "Time window of 10 seconds",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  ) 

```

```{r 20s Window}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Visualisation of the static events of all tracks with a time window of 20 seconds."
#| label: fig-20-window

# Plot the static events

theme_x_cord <- list(
  theme_minimal(),
  coord_sf(datum = 2056),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
    plot.title = element_text(size = 10, face = "bold")
  )
)

# hide legend for the first 5 plots to prevent duplicates
hide_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  guide = "none"
)

# show full legend on the very last plot, forcing all 3 colors to display
show_full_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  na.value = "grey",
  limits = c(TRUE, FALSE, NA), # <--- Keep this order
  labels = c("Static", "Moving", "Unknown"), # <--- Match the order of limits exactly without names
  name = "Status"
)

p1 <- ggplot(Strava_1_processed_20 ) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 1") +
  theme_x_cord

p2 <- ggplot(Strava_2_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 2") +
  theme_x_cord

p3 <- ggplot(Swisstopo_1_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 1") +
  theme_x_cord

p4 <- ggplot(Swisstopo_2_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 2") +
  theme_x_cord

p5 <- ggplot(Posmo_1_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Posmo 1") +
  theme_x_cord

p6 <- ggplot(Posmo_2_processed_20) +
  geom_sf(aes(color = is_static), size = 1) +
  show_full_legend +
  labs(title = "Posmo 2") +
  theme_x_cord

# creates a grid layout with 3 columns, and collects the legends into one clean spot
p1 + p2 + p3 + p4 + p5 + p6 + 
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Static vs. Moving",
    subtitle = "Time window of 20 seconds",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  ) 

```

```{r 30s Window}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Visualisation of the static events of all tracks with a time window of 30 seconds."
#| label: fig-30-window

# Plot the static events

theme_x_cord <- list(
  theme_minimal(),
  coord_sf(datum = 2056),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
    plot.title = element_text(size = 10, face = "bold")
  )
)

# hide legend for the first 5 plots to prevent duplicates
hide_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  guide = "none"
)

# show full legend on the very last plot, forcing all 3 colors to display
show_full_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  limits = c(TRUE, FALSE), # <--- Keep this order
  labels = c("Static", "Moving"), # <--- Match the order of limits exactly without names
  name = "Status"
)

p1 <- ggplot(Strava_1_processed_30 ) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 1") +
  theme_x_cord

p2 <- ggplot(Strava_2_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 2") +
  theme_x_cord

p3 <- ggplot(Swisstopo_1_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 1") +
  theme_x_cord

p4 <- ggplot(Swisstopo_2_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 2") +
  theme_x_cord

p5 <- ggplot(Posmo_1_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Posmo 1") +
  theme_x_cord

p6 <- ggplot(Posmo_2_processed_30) +
  geom_sf(aes(color = is_static), size = 1) +
  show_full_legend +
  labs(title = "Posmo 2") +
  theme_x_cord

# creates a grid layout with 3 columns, and collects the legends into one clean spot
p1 + p2 + p3 + p4 + p5 + p6 + 
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Static vs. Moving",
    subtitle = "Time window of 30 seconds",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  ) 

```

```{r 40s Window}
#| fig-width: 10
#| fig-height: 11
#| fig-cap: "Visualisation of the static events of all tracks with a time window of 40 seconds."
#| label: fig-40-window

# Plot the static events

theme_x_cord <- list(
  theme_minimal(),
  coord_sf(datum = 2056),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
    plot.title = element_text(size = 10, face = "bold")
  )
)

# hide legend for the first 5 plots to prevent duplicates
hide_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  guide = "none"
)

# show full legend on the very last plot, forcing all 3 colors to display
show_full_legend <- scale_color_manual(
  values = c("TRUE" = "darkgreen", "FALSE" = "#ADD8E666"),
  limits = c(TRUE, FALSE), # <--- Keep this order
  labels = c("Static", "Moving"), # <--- Match the order of limits exactly without names
  name = "Status"
)

p1 <- ggplot(Strava_1_processed_40 ) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 1") +
  theme_x_cord

p2 <- ggplot(Strava_2_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Strava 2") +
  theme_x_cord

p3 <- ggplot(Swisstopo_1_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 1") +
  theme_x_cord

p4 <- ggplot(Swisstopo_2_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Swisstopo 2") +
  theme_x_cord

p5 <- ggplot(Posmo_1_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  hide_legend +
  labs(title = "Posmo 1") +
  theme_x_cord

p6 <- ggplot(Posmo_2_processed_40) +
  geom_sf(aes(color = is_static), size = 1) +
  show_full_legend +
  labs(title = "Posmo 2") +
  theme_x_cord

# creates a grid layout with 3 columns, and collects the legends into one clean spot
p1 + p2 + p3 + p4 + p5 + p6 + 
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Static vs. Moving",
    subtitle = "Time window of 40 seconds",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  ) 

```

@fig-10s-strava2 and @fig-30s-posmo1 show two examples for varying time windows and mobile apps of the visualised stop detection inside different buffer sizes computed with different cluster distances for the static events. 

```{r Example: Strava 2 10s Window}
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Visualisation of the static events (with a time window of 10 seconds) inside the buffer for different cluster distances and buffer sizes for strava 2."
#| label: fig-10s-strava2


# plot the combinations of clustering and buffering around waypoint

theme_x_cord <- list(
  coord_sf(datum = 2056),
  theme_minimal(),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
    plot.title = element_text(size = 9, face = "bold"),
    axis.title = element_blank()
  )
)

plot_liste <- list()

for (md in max_dist_liste) {
  
  filtered <- Strava_2_10_stops_list |> 
    filter(max_dist_setting == md)
  
  for (buf in wp_buffer_liste) {
    
    current_col <- paste0("in_buf_md", md, "_b", buf)
    
    p <- ggplot() +
      geom_sf(data = st_buffer(wp_olten_clean, buf), 
              fill = "gray70", 
              alpha = 0.50, 
              color = NA) +
      geom_sf(data = filtered, 
              aes(color = as.character(.data[[current_col]])), 
              size = 1) +
      scale_color_manual(
        values = c("1" = "darkgreen", "0" = "blue"),
        labels = c("1" = "Hit (TP)", "0" = "False (FP)"),
        drop = FALSE
      ) +
      labs(title = paste0("MD: ", md, "m | Buffer: ", buf, "m"), 
           color = "Inside of the buffer") +
      theme_x_cord
    
    plot_liste[[paste0("md", md, "_b", buf)]] <- p
  }
}

wrap_plots(plot_liste, ncol = 3) + 
  plot_layout(guides = "collect") + 
  plot_annotation(
    title = "Strava 2 - 9 combinations of clustering and buffering around waypoints",
    subtitle = "Temporal window for static events = 10 seconds | Cluster-Distanz (MD) | Waypoint-Buffer (Buffer)"
  )
```

```{r Example: Posmo 1 30s Window}
#| fig-width: 10
#| fig-cap: "Visualisation of the static events (with a time window of 30 seconds) inside the buffer for different cluster distances and buffer sizes for Posmo 1"
#| label: fig-30s-posmo1


# plot the combinations of clustering and buffering around waypoint

theme_x_cord <- list(
  coord_sf(datum = 2056),
  theme_minimal(),
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
    plot.title = element_text(size = 9, face = "bold"),
    axis.title = element_blank()
  )
)

plot_liste <- list()

for (md in max_dist_liste) {
  
  filtered <- Posmo_1_30_stops_list |> 
    filter(max_dist_setting == md)
  
  for (buf in wp_buffer_liste) {
    
    current_col <- paste0("in_buf_md", md, "_b", buf)
    
    p <- ggplot() +
      geom_sf(data = st_buffer(wp_wollerau_clean, buf), 
              fill = "gray70", 
              alpha = 0.50, 
              color = NA) +
      geom_sf(data = filtered, 
              aes(color = as.character(.data[[current_col]])), 
              size = 1) +
      scale_color_manual(
        values = c("1" = "darkgreen", "0" = "blue"),
        labels = c("1" = "Hit (TP)", "0" = "Outside of Buffer (FP)"),
        drop = FALSE
      ) +
      labs(title = paste0("MD: ", md, "m | Buffer: ", buf, "m"), 
           color = "Inside of the buffer") +
      theme_x_cord
    
    plot_liste[[paste0("md", md, "_b", buf)]] <- p
  }
}

wrap_plots(plot_liste, ncol = 3) + 
  plot_layout(guides = "collect") + 
  plot_annotation(
    title = "Posmo 1 - 9 combinations of clustering and buffering around waypoints",
    subtitle = "Temporal window for static events = 30 seconds | Cluster-Distanz (MD) | Waypoint-Buffer (Buffer)"
  )
```

To compare the effectiveness of using different time windows to classify static events, various clustering distances to create single stops and different buffers around the waypoints @tbl-parameters-all-prov compares the performance of the different parameters for all three mobile apps together. The table shows for each parameter combination the total stops created (Total Stops Found),  stops correctly identified inside the buffer (TP), the stops identified outside of the buffer (FP) and the number of buffers for which no stops were detected (FN). Furthermore, the performance metrics precision ($\frac{TP}{TP+FP}$) and recall ($\frac{TP}{TP+FN}$) are calculated in the table and recall is used to highlight the 5 best (green) and 5 worst (red) performances. Overall 5 metre waypoint buffers detected the fewest stops inside the buffers and performed the worst. Waypoint buffers of 15 metres in connection with big temporal windows performed the best in detecting true stops as well as only having few false stops. 

```{r Parameter Table with Highlighting}
#| tbl-cap: "Parameter Table evaluating out of maximum 88 possible ground truth waypoint targets across all combined tracking apps. The top 5 performing rows for recall values are highlighted in green and the bottom 5 in red."
#| label: tbl-parameters-all-prov

# helper functions for dynamic column evaluation
extract_sum <- function(df, md, buf, val) {
  col_name <- paste0("in_buf_md", md, "_b", buf)
  if (!col_name %in% colnames(df)) return(0)
  sum(df[[col_name]] == val, na.rm = TRUE)
}

# function to process a dataset
get_all_permutations <- function(df, app_name, track_id, window_sec, total_gt) {
  if (nrow(df) == 0) return(data.frame())
  distances <- c(5, 10, 15)
  buffers   <- c(5, 10, 15)
  param_grid <- expand.grid(Max_Dist = distances, WP_Buffer = buffers)
  
  map_dfr(1:nrow(param_grid), function(i) {
    md  <- param_grid$Max_Dist[i]
    buf <- param_grid$WP_Buffer[i]
    df_sub <- df |> 
      filter(max_dist_setting == md)
    
    if (nrow(df_sub) == 0) return(data.frame())
    
    data.frame(
      App         = app_name,
      Track       = track_id,
      Window      = paste0(window_sec, "s"),
      Max_Dist    = md,
      WP_Buffer   = buf,
      Total_Stops = nrow(df_sub),
      True_Pos    = extract_sum(df_sub, md, buf, TRUE),
      False_Pos   = extract_sum(df_sub, md, buf, FALSE),
      Total_GT    = total_gt
    )
  })
}

# explicit ground truth variables
gt_strava_t1 <- nrow(wp_olten_clean)    # 14
gt_strava_t2 <- nrow(wp_olten_clean)    # 14
gt_swiss_t1  <- nrow(wp_wollerau_clean) # 15
gt_swiss_t2  <- nrow(wp_wollerau_clean) # 15
gt_posmo_t1  <- nrow(wp_wollerau_clean) # 15
gt_posmo_t2  <- nrow(wp_wollerau_clean) # 15

# maximum possible waypoints to detect across all 6 tracks combined
MAX_GLOBAL_GT <- (14 * 2) + (15 * 4) 

# run calculations over the 24 clean datasets
master_raw_metrics <- bind_rows(
  get_all_permutations(Strava_1_10_stops_list, "Strava", "T1", 10, gt_strava_t1),
  get_all_permutations(Strava_2_10_stops_list, "Strava", "T2", 10, gt_strava_t2),
  get_all_permutations(Swisstopo_1_10_stops_list, "Swisstopo", "T1", 10, gt_swiss_t1),
  get_all_permutations(Swisstopo_2_10_stops_list, "Swisstopo", "T2", 10, gt_swiss_t2),
  get_all_permutations(Posmo_1_10_stops_list, "Posmo", "T1", 10, gt_posmo_t1),
  get_all_permutations(Posmo_2_10_stops_list, "Posmo", "T2", 10, gt_posmo_t2),
  
  get_all_permutations(Strava_1_20_stops_list, "Strava", "T1", 20, gt_strava_t1),
  get_all_permutations(Strava_2_20_stops_list, "Strava", "T2", 20, gt_strava_t2),
  get_all_permutations(Swisstopo_1_20_stops_list, "Swisstopo", "T1", 20, gt_swiss_t1),
  get_all_permutations(Swisstopo_2_20_stops_list, "Swisstopo", "T2", 20, gt_swiss_t2),
  get_all_permutations(Posmo_1_20_stops_list, "Posmo", "T1", 20, gt_posmo_t1),
  get_all_permutations(Posmo_2_20_stops_list, "Posmo", "T2", 20, gt_posmo_t2),
  
  get_all_permutations(Strava_1_30_stops_list, "Strava", "T1", 30, gt_strava_t1),
  get_all_permutations(Strava_2_30_stops_list, "Strava", "T2", 30, gt_strava_t2),
  get_all_permutations(Swisstopo_1_30_stops_list, "Swisstopo", "T1", 30, gt_swiss_t1),
  get_all_permutations(Swisstopo_2_30_stops_list, "Swisstopo", "T2", 30, gt_swiss_t2),
  get_all_permutations(Posmo_1_30_stops_list, "Posmo", "T1", 30, gt_posmo_t1),
  get_all_permutations(Posmo_2_30_stops_list, "Posmo", "T2", 30, gt_posmo_t2),
  
  get_all_permutations(Strava_1_40_stops_list, "Strava", "T1", 40, gt_strava_t1),
  get_all_permutations(Strava_2_40_stops_list, "Strava", "T2", 40, gt_strava_t2),
  get_all_permutations(Swisstopo_1_40_stops_list, "Swisstopo", "T1", 40, gt_swiss_t1),
  get_all_permutations(Swisstopo_2_40_stops_list, "Swisstopo", "T2", 40, gt_swiss_t2),
  get_all_permutations(Posmo_1_40_stops_list, "Posmo", "T1", 40, gt_posmo_t1),
  get_all_permutations(Posmo_2_40_stops_list, "Posmo", "T2", 40, gt_posmo_t2)
)

# corrected mathematical aggregation
global_parameter_table <- master_raw_metrics |>
  group_by(Window, Max_Dist, WP_Buffer) |>
  summarise(
    Total_Stops_Found  = sum(Total_Stops),
    True_Positives     = sum(True_Pos),
    False_Positives    = sum(False_Pos),
    
    # False Negatives = Total available ground truth points (88) minus the ones we successfully hit
    False_Negatives    = MAX_GLOBAL_GT - True_Positives,
    
    # Precision and recall 
    Detection_Precision = round((True_Positives / (True_Positives + False_Positives)) * 100, 1),
    Detection_Recall = round((True_Positives / MAX_GLOBAL_GT) * 100, 1),
    .groups = "drop"
  ) |>
  arrange(Window, Max_Dist, WP_Buffer)

# rank performances
ranked_indices <- global_parameter_table |>
  mutate(row_id = row_number()) |>
  arrange(desc(Detection_Recall), False_Positives)

top_5_rows    <- head(ranked_indices$row_id, 5)
bottom_5_rows <- tail(ranked_indices$row_id, 5)

# render table
global_parameter_table |>
  select(Window, Max_Dist, WP_Buffer, Total_Stops_Found, True_Positives, False_Positives, False_Negatives, Detection_Precision, Detection_Recall) |>
  knitr::kable(
    format = "html",
    col.names = c("Time Window", "Maximal Clustering Distance (m)", "Waypoint Buffer (m)", "Total Stops Found", "True Positives (TP)", "False Positives (FP)", "False Negatives (FN)", "Precision (%)", "Recall (%)")
  ) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  row_spec(top_5_rows, background = "#d4edda", extra_css = "font-weight: bold; color: #155724;") |>
  row_spec(bottom_5_rows, background = "#f8d7da", extra_css = "color: #721c24;")
```

Form the precision and recall values we can calculate The F1-score $F_1 = 2 \times \frac{\text{Precision} \times \text{Recall}}{\text{Precision} + \text{Recall}}$ which is a further indication of performance. @fig-F1-together shows this F1-score for all the apps together and @fig-F1-by-App shows the F1-scores more precisely for each app. @fig-F1-together underlines the results of @tbl-parameters-all-prov that for all apps combined the 5 metre waypoint buffer performs the worst in and especially in connection with a 10 seconds time window to detect static events. The figure further highlights that the 15 metres waypoint buffer in connection with a big temporal window for static events and rather higher maximal clustering distances perform the best.

```{r F1 score for all}
#| fig-width: 8
#| fig-cap: "Visualisation of F1 score for the different parameters across all combined tracking apps."
#| label: fig-F1-together

  
global_parameter_table <- global_parameter_table |>
  mutate(
    Precision = True_Positives / (True_Positives + False_Positives),
    Recall    = True_Positives / (True_Positives + False_Negatives),
    
    F1 = ifelse(
      (Precision + Recall) == 0,
      0,
      2 * (Precision * Recall) / (Precision + Recall)
    )
  )


ggplot(global_parameter_table, aes(x = Max_Dist, y = WP_Buffer, fill = F1)) +
  geom_tile(color = "black", linewidth = 0.1) +
  facet_wrap(~ Window) +
  scale_fill_gradient(low = "red", high = "blue", limits = c(0, 1)) +
  labs(title = "F1 Score across Parameters", 
       x = "Maximal Clustering Distance (m)",
       y = "Waypoint Buffer Size (m)",
       fill = "F1")

```

@fig-F1-by-App indicates that the Strava app performs overall the best where as Posmo on the other hand performs very bad especially for the 10 seconds time window. This is also further highlighted in @tbl-app-performance. 

```{r F1 score by App}
#| fig-width: 8
#| fig-cap: "Visualisation of F1 score for the different parameters for all apps seperatly."
#| label: fig-F1-by-App

add_metrics <- function(df) {
  df |>
    mutate(
      Precision = True_Pos / (True_Pos + False_Pos),
      Recall    = True_Pos / Total_GT,
      
      F1 = ifelse(
        (Precision + Recall) == 0,
        0,
        2 * Precision * Recall / (Precision + Recall)
      )
    )
}
master_apps <- add_metrics(master_raw_metrics)

ggplot(master_apps, aes(x = Max_Dist, y = WP_Buffer, fill = F1)) +
  geom_tile(color = "black", linewidth = 0.1) +
  facet_grid(App ~ Window) +
  scale_fill_gradient(low = "red", high = "blue", limits = c(0, 1)) +
  labs(
    title = "F1 Score across Parameters (by App)",
    x = "Maximal Clustering Distance (m)",
    y = "Waypoint Buffer Size (m)",
    fill = "F1"
  ) 

```

```{r Mean precision and recall metrics}
#| tbl-cap: "Mean precision and recall metrics grouped by tracking application and time window settings."
#| label: tbl-app-performance

result_all <- master_apps  |> 
  group_by(App, Window) |>
  summarise(
    precision_mean = round(mean(Precision, na.rm = TRUE), 3),
    recall_mean = round(mean(Recall, na.rm = TRUE), 3),
    F1_mean = round(mean(F1, na.rm = TRUE), 3),
    .groups = "drop"
  ) |>
  arrange(App, Window)

result_all |>
  knitr::kable(
    format = "html",
    col.names = c("Application", "Time Window", "Mean Precision", "Mean Recall", "Mean F1-score"),
    align = c("l", "c", "r", "r") # Left, Center, Right, Right alignment
  ) |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE
  )

```

However the best F1-scores for Posmo and Swisstopo were not achieved for the 30 seconds time window but rather the 30 seconds time window with a 10 metre clustering distance (@tbl-app-best-performance). 

```{r Top performance for each app}
#| tbl-cap: "Summary of F1-Scores per tracking app, showing the best performance alongside its optimal parameters, dataset averages and the amount of perfect performances per app. Strava shows multiple top performanece and the Best F1-score therefore only shows one example of a possible parameter combination for a perfect F1-score."
#| label: tbl-app-best-performance

app_summary <- master_apps |>
  group_by(App) |>
  summarise(
    # Find the index of the highest F1 score for this app
    best_idx = which.max(F1),
    
    # Extract the highest F1 value
    max_val  = round(F1[best_idx], 3),
    
    # Extract the matching parameter values
    win    = Window[best_idx],
    b_dist   = Max_Dist[best_idx],
    b_buf    = WP_Buffer[best_idx],
    
    # Combine them into a clean string: "Value (Window, Dist, Buffer)"
    Best_F1  = paste0(max_val, " (", win, ", ", b_dist, "m, ", b_buf, "m)"),
    
    # Calculate the remaining metrics
    Mean_F1          = round(mean(F1, na.rm = TRUE), 3),
    Perfect_Score_N  = sum(F1 == 1, na.rm = TRUE),
    .groups = "drop"
  ) |>
  # Drop the temporary helper columns used for extraction
  select(App, Best_F1, Mean_F1, Perfect_Score_N)

app_summary |>
  knitr::kable(
    format = "html",
    col.names = c("Application", "Best F1-Score (Time Window	Clustering, Maximal Clustering Distance, Waypoint Buffer)", "Mean F1-Score", "Perfect Setups (F1 = 1)"),
    align = c("l", "l", "r", "c", "c") # Changed second column to left-align for text readability
  ) |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )
```

Grouping the results for the different parameters separately it is further highlighted that the 30 and 40 seconds windows create the best F-1 scores (@tbl-best-perf-tw). The different clustering distances on the other hand seem to not have a big influence on the F1-score where as the buffer size of 5 metres results in significantly lower overall F1-scores as already mentioned abpve (@tbl-best-perf-CD and @tbl-best-perf-buff). 

::: {#tbl-panel layout-ncol=3}

```{r}
#| tbl-cap: "Average perfromance for each time window."
#| label: tbl-best-perf-tw

master_apps  |> 
  group_by(Window) |>
  summarise(
    mean_F1 = mean(F1, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  knitr::kable(format = "html", col.names = c("Window", "Mean F1"))
```

```{r}
#| tbl-cap: "Avereage perfrormance for each clustering distance."
#| label: tbl-best-perf-CD

master_apps |>
  group_by(Max_Dist) |>
  summarise(
    mean_F1 = mean(F1, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  knitr::kable(format = "html", col.names = c("Max Dist (m)", "Mean F1"))
```

```{r}
#| tbl-cap: "Avereage perfrormance for each waypoint buffer size."
#| label: tbl-best-perf-buff

master_apps |>
  group_by(WP_Buffer) |>
  summarise(
    mean_F1 = mean(F1, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  knitr::kable(format = "html", col.names = c("Buffer (m)", "Mean F1"))

```

Top performance for all the different parameters separately.
:::


## Discussion

The results of the analysis are based on F1 scores. This metric was chosen because it incorporates all three of our analysis parameters: TP, FP, and FN. The F1 score is calculated using precision and recall metrics. The precision value compares the TP with the FP—that is, the ratio of correctly to incorrectly identified stops—while recall calculates a value that indicates whether all ground truth stops were identified. The closer these parameter values are to 1, the more accurate the method. The F1 score combines both parameters and yields a value between 0 and 1. Thus, the closer the F1 score value is to 1, the more accurate the method. At first glance, the segmentation method appears to show high F1 scores, though with a few exceptions.

In @fig-F1-together, the low F1 score indicates that the 5-meter buffer is a poor choice. The reason for this is that the stops usually covered a large area, and the exercises required a somewhat larger radius. Likewise, GPS inaccuracies are more sensitive in a small buffer and can deviate much more quickly. This negative behavior of the 5m buffer is also evident in @tbl-best-perf-buff. It is also visible that the 10-second intervals yield significantly lower values. However, the reason for this is better shown in @fig-F1-by-App. The Posmo app delivers very low values at short time intervals. Even at 20-second intervals, the values are still not outstanding. This is a weakness of the chosen methodology, since Posmo provides a GPS signal on average every 15 seconds and there are not enough neighbors within the 10-second interval. This results in a very large number of NA stepMean values. Even in the 20-second intervals, there are still some NA stepMean values, since on average only 1 to 2 neighbors are detected within that time window. Even with 40-second intervals, there are only an average of 3 neighbors. Strava, on the other hand, sends a signal every 2 seconds, which means that in a 10-second interval, the stepMean is already calculated from 5 neighbors; with Swisstopo, this happens every second, allowing stepMean to be calculated from 10 neighbors.

One possible solution to this problem is to use interpolation over a specific time interval, rather than calculating a stepMean based on existing values within a time interval. Interpolation makes it possible to create comparable time intervals across all apps, with coordinates being linearly interpolated and adjusted. The disadvantage is that, with long signal intervals such as in Posmo, interpolation can lead to inaccurate coordinate shifts. Interpolation was also the method originally chosen, but after several failed attempts, it was identified as a limitation of this study. Consequently, due to the alternative methodology, the results from apps like Posmo—which transmit signals at 15-second intervals—should be interpreted with caution or are in some cases meaningless, especially when a time interval contains an average of 0 to 1 data points.

@tbl-parameters-all-prov shows that both precision and recall improve with a longer time interval. @tbl-app-performance also reveals the app’s impact, showing that Posmo is significantly negatively affected by recall. This means that in 10-second intervals, the identification of all 15 existing waypoint buffers in Wollerau is very inaccurate, as it is complicated by the NA values. In this study, not only does recall have a negative impact on the F1 score, but precision does as well. In @tbl-app-performance, Swisstopo yields lower F1 scores at shorter time intervals. Recall remains equally good across time intervals, but precision varies. The smaller the time interval, the lower the F1 score, as fewer neighbors are available for determining correct stops, and the risk of identifying incorrect stops increases. This is not only the case for Swisstopo; according to @tbl-best-perf-tw, this behavior holds true regardless of the app.

It is particularly interesting that the waypoints were created using Swisstopo—which has the best sampling interval—yet Strava still delivers better F1 scores than Swisstopo in @tbl-app-best-performance. Strava even manages to achieve a perfect F1 score of 1 in 20 different combinations. In contrast, Posmo and Swisstopo never reach 1, with Swisstopo’s best value being 0.966. Since Strava was the only app used in Olten, it is possible that the running behavior in Olten was more suitable for the segmentation method—for example, longer breaks were taken during stops, or no interruptions occurred during the run, or just the App which was indeed the case. The only parameter that makes hardly any difference to the results in @tbl-best-perf-CD is the cluster selection. This is because FP data points are either very close to each other or far apart anyway. For this to make a difference, the cluster selection would have to be taken to the extreme.


##conclusion

- je länger time window, desto mehr nachbarn können verwendet werden, um einen guten wert zu bekommen
- 5m schlecht
- posmo schlecht, vor allem bei 10 und 20 sek, wobei interpolation helfen könnte
- Strava sehr gut 

## Appendix

### AI use

### Wordcount

<!-- after installing the wordcountadding, remove the line "#| eval: false" -->

```{r}
# #| eval: false
# library("pacman")
# p_install_gh("benmarwick/wordcountaddin")

wordcountaddin::word_count("index.qmd")
```


